---
title: "Dataframes - In Class lab"
author: "Dr. Shoemaker"
date: "DATA 2401"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Working with built-in data sets

Load R's "USPersonalExpenditure" dataset using the `data()` function

This will produce a data frame called `USPersonalExpenditure`
```{r}
data("USPersonalExpenditure")
```

The variable `USPersonalExpenditure` is now accessible to you. Unfortunately, it's not a data frame (it's actually a matrix, you can find that out using) 

Test this using the `is.data.frame()` function:

```{r}
is.data.frame(USPersonalExpenditure)
```

Luckily, you can pass the USPersonalExpenditure variable as an argument to the `data.frame()` function to convert it a data frame.

Do this, storing the result in a new variable

```{r}
us_spend <- data.frame(USPersonalExpenditure)
```


What are the current column names of your dataframe?
```{r}
colnames(us_spend)
```


Consider: why are they so strange? (Think about whether you could use a number like 1940 with dollar notation!)

What are the current row names of your dataframe?
```{r}
rownames(us_spend)
```


Add a column "Category" to your data frame that contains the rownames

```{r}
us_spend['category'] <- rownames(us_spend)

head(us_spend)
```

Consider how this data isn't tidy. Use the `pivot_longer()` function from last class to make the data tidy. 

```{r}

us <- pivot_longer(data = us_spend, 
             cols = c(`X1940`, `X1945`,`X1950`,`X1955`,`X1960`), 
             names_to = "year", 
             values_to = "spent")

us
```


How much money was spent on personal care in 1940?
```{r}
cat <- us$category
year<- us$year
spent <- us$spent
cat


spent[cat == 'Personal Care'[year == 'X1940']]




```


How much money was spent on Food and Tobacco in 1960?
```{r}
spent[cat=='Food and Tobacco'[year == 'X1960']]

```


What was the highest expenditure category in 1960?
```{r}

cat[spent == max(spent[year == 'X1960'])]

```


Define a function `lowest_category` that takes in a year as a parameter, and returns the lowest spending category of that year
```{r}
lowest_category <- cat[spent == min(spent[year == min(year)])]
lowest_category

```

```{r}

```


Using your function, determine the lowest spending category of each year
Hint: use the `sapply()` function to apply your function to a vector of years

```{r}

```
